---
title: "sql-workshop"
author: "Charles Lang"
output: html_document
---

Before you follow the directions below, please take a screenshot of your AWS console showing the running database and upload it to your repo.

## Connect to AWS MySQL Database
```{r}
#install.packages("DBI", "RMySQL")

library(DBI)
library(RMySQL)
```
```{r}

db_user <- 'admin'
db_password <- 'testsql!'
db_name <- 'oudb'
db_host <- 'database-1.crktxevex7kw.us-east-2.rds.amazonaws.com'
db_port <- 3306
```
```{r}

mydb <- dbConnect(MySQL(), user = db_user, password = db_password, dbname = db_name, host = db_host, port = db_port)
summary(mydb)

```

## Load OU Data
```{r}
#Student demographic data
studentInfo <- read.csv("studentInfo.csv", header = TRUE)
#Student assessment data
studentAssessment <- read.csv("studentAssessment.csv", header = TRUE)
#Course data
courses <- read.csv("courses.csv", header = TRUE)
studentRegistration <- read.csv("studentRegistration.csv", header = TRUE)
```

## Write data to the DB using the DBI package
```{r}
#List the tables in the DB - should be zero
dbListTables(mydb)
```
```{r}
#Write a new table to the DB
dbWriteTable(mydb, "studentInfo", studentInfo)
dbWriteTable(mydb, "studentAssessment", studentAssessment)
dbWriteTable(mydb, "courses", courses)
dbWriteTable(mydb, "studentRegistration", studentRegistration)
```
```{r}
#List tables to see that table was added
dbListTables(mydb)

#Read a particular table
dbReadTable(mydb, 'studentInfo')
```
```{r}
#EXERCISE 1
#Make two toy data sets (two dataframes?so two tables?)
#with at least three variables and at least 30 rows each in them(manually?). 
#Have a mix of numeric and character variables. 
TeacherFirstName <- c(rep("John",10),rep("Bob",10),rep("Rachel",10))
TeacherLastName <- c(rep("A",10),rep("B",10),rep("C",10))
TeacherAge <-c(rep(25,10),rep(30,10),rep(40,10))
Teacher_DF <-data.frame(TeacherFirstName,TeacherLastName,TeacherAge)
Teacher_DF
```
```{r}
SchoolName <-c(rep("a",10),rep("b",10),rep("c",10))
SchoolNationalRanking <-c(floor(runif(30,min=50,max=101)))
SchoolStateRanking <-c(floor(runif(30,min=1,max=31)))
School_DF <-data.frame(SchoolName,SchoolNationalRanking,SchoolStateRanking)
School_DF
```
#```{r}
#dbRemoveTable(mydb,"Teachers")
#```


```{r}
#Transfer these dataframes to your SQL database using the DBI commands. Name the tables whatever you like.
dbWriteTable(mydb, "Teachers", Teacher_DF)
```

#```{r}
#dbRemoveTable(mydb,"SchoolRankings")
#```

```{r}
dbWriteTable(mydb,"SchoolRankings",School_DF)
```

```{r}
#test if the table successfully added 
dbListTables(mydb)
```


## Getting into SQL - READING
```{r}
#Query a portion of the database (always returns dataframe)
dbGetQuery(mydb, "SELECT * FROM studentInfo LIMIT 10;")

dbGetQuery(mydb, "SELECT * FROM studentInfo ORDER BY id_student LIMIT 10;")

dbGetQuery(mydb, "SELECT id_student, gender FROM studentInfo ORDER BY id_student DESC LIMIT 10;") #Order listed will be reflected in order in table

dbGetQuery(mydb, "SELECT id_student AS 'Student ID', gender FROM studentInfo LIMIT 10;") #SQL Standard says quotes for literal strings and double quotes for everything else but that conflicts with R

#Count the number of rows
dbGetQuery(mydb, "SELECT COUNT(*) FROM studentAssessment;")

#Using a WHERE statement on all columns
dbGetQuery(mydb, "SELECT COUNT(*) FROM studentAssessment WHERE score > 50;")

#Using a WHERE statement on a single column (will not include missing data)
dbGetQuery(mydb, "SELECT COUNT(score) FROM studentAssessment WHERE score > 50;")

#Using an AND statement
dbGetQuery(mydb, "SELECT COUNT(*) FROM studentAssessment WHERE score > 50 AND id_assessment = '1752';")

#dbGetQuery(mydb,"SELECT COUNT(*) FROM studentAssessment WHERE id_assessment = '1752';")

```
```{r}
#EXERCISE 2
#Read one of your toy data tables, make sure the output is ordered in descending order, you rename one of the variables and the output is limited to the first 20 rows.
dbGetQuery(mydb,"SELECT * FROM Teachers ORDER BY TeacherAge DESC LIMIT 20;")
#Read the other table according to a condition of one of the variables.
#read top 10 schools of the state in ascending order from the schoolranking table (by default, ascending order)
dbGetQuery(mydb, "SELECT * FROM SchoolRankings WHERE SchoolStateRanking <11 ORDER BY SchoolStateRanking;")
```


## Getting into SQL - UPDATING

```{r}
#Count rows
dbGetQuery(mydb, "SELECT COUNT(*) FROM studentAssessment;")

#Add a row
dbGetQuery(mydb, "INSERT INTO studentAssessment (id_assessment, id_student, date_submitted, is_banked, score) VALUES ('00001', '1', '20', '0', '50');")

#Count rows again
dbGetQuery(mydb, "SELECT COUNT(*) FROM studentAssessment;")

#View inserted row
dbGetQuery(mydb, "SELECT * FROM studentAssessment ORDER BY id_student LIMIT 10;")

#Add a row with missing values
dbGetQuery(mydb, "INSERT INTO studentAssessment (id_assessment, id_student, date_submitted) VALUES ('00001', '1', '20');")

#View inserted row
dbGetQuery(mydb, "SELECT * FROM studentAssessment ORDER BY id_student LIMIT 10;")

#Update a row
dbGetQuery(mydb, "UPDATE studentAssessment SET score = '20' WHERE id_student = 1;")

dbGetQuery(mydb, "SELECT id_student, score FROM studentAssessment ORDER BY id_student LIMIT 10;")

#Update a row with NULL
dbGetQuery(mydb, "UPDATE studentAssessment SET score = 'NULL' WHERE id_student = 6516;")

#Delete a row (destructive)
dbGetQuery(mydb, "DELETE FROM studentAssessment WHERE id_student = 1;")

dbGetQuery(mydb, "SELECT * FROM studentAssessment ORDER BY id_student LIMIT 10;")


dbGetQuery(mydb, "SELECT COUNT(*) FROM studentAssessment;")
```

## Add/Deleting Table


```{r}
#Creating a new table in SQL
dbGetQuery(mydb,"CREATE TABLE test (
  score INTEGER, 
  student TEXT
  );")

dbListTables(mydb)

#Inserting data into the table
dbGetQuery(mydb, "INSERT INTO test VALUES ( 10, 'Amy' );")
dbGetQuery(mydb, "INSERT INTO test VALUES ( 11, 'Jen' );")
dbGetQuery(mydb, "INSERT INTO test VALUES ( 9, 'Frank' );")

dbGetQuery(mydb, "SELECT * FROM test;")

#Inserting a NULL row
dbGetQuery(mydb, "INSERT INTO test DEFAULT VALUES;") #Will not work use instead:

dbGetQuery(mydb,"INSERT INTO test (score, student) SELECT score, id_student FROM studentAssessment;")

#Delete a table
dbGetQuery(mydb, "DROP TABLE test;")

dbGetQuery(mydb, "SELECT * FROM test;") #This should produce an error since your table no longer exists

#Delete a table if it exists
dbGetQuery(mydb, "DROP TABLE IF EXISTS test;") #No error since it is only if it exists

```
```{r}
#EXERCISE 4
#Create a table that is exactly the same as your first toy data table but this time use SQL commands. Display your new table. Then delete the original table. 
#delete the table if it's already existed 
dbGetQuery(mydb, "DROP TABLE IF EXISTS NewTeachers;") 
dbGetQuery(mydb,"CREATE TABLE NewTeachers (
           TeacherFirstName TEXT,
           TeacherLastName TEXT,
           TeacherAge INTEGER
           );")
dbListTables(mydb)
#add data from the teachers table (ONLY can be run one time, otherwise it'll add up)
dbGetQuery(mydb, "INSERT INTO NewTeachers (TeacherFirstName,TeacherLastName, TeacherAge) SELECT TeacherFirstName, TeacherLastName, TeacherAge FROM Teachers ;")
#CHECK if the data is inserted correctly
dbGetQuery(mydb,"SELECT * FROM NewTeachers;")
#insert new rows
#dbGetQuery(mydb, "INSERT INTO NewTeachers VALUES('John', 'A',25),('Bob','B',30),('Rachel','C',40);")
#drop the original table (i.e., Teachers) by using DBI
#dbRemoveTable(mydb,"Teachers")
#dbGetQuery(mydb, "DROP TABLE Teachers;")
dbGetQuery(mydb, "DROP TABLE IF EXISTS Teachers;")
#check if the table "teachers" has been successfully dropped
dbListTables(mydb)
```

# NULL Value
```{r}
#NULL is a state (similar to R), represents the lack of a value. But is not compatible with R backend so this code doesn't work as part of dbGetQuery()

#This doesn't work because NULL is not a value
SELECT * FROM test WHERE score = NULL;

#Instead use
SELECT * FROM test WHERE score is NULL;

```

# Constraints
```{r}
#Create table where student column *cannot* be NULL
dbGetQuery(mydb,"CREATE TABLE test2 (
  score INTEGER, 
  student TEXT NOT NULL
  );")

dbGetQuery(mydb, "DROP TABLE IF EXISTS test2;")

#Create table where score column *cannot* be NULL
dbGetQuery(mydb,"CREATE TABLE test2 (
   score INTEGER DEFAULT 0, 
   student TEXT
   );")

dbGetQuery(mydb,"INSERT INTO test2 (score, student) VALUES ('1', 'A');")
dbGetQuery(mydb,"INSERT INTO test2 (student) VALUES ('B');")

dbGetQuery(mydb, "SELECT * FROM test2;")

dbGetQuery(mydb, "DROP TABLE IF EXISTS test2;")

dbGetQuery(mydb,"CREATE TABLE test2 (
  score INTEGER UNIQUE, 
  student TEXT
  );")

dbGetQuery(mydb,"INSERT INTO test2 (score, student) VALUES ('1', 'A');")

#Error because of unique
dbGetQuery(mydb,"INSERT INTO test2 (score, student) VALUES ('1', 'A');")

#NULL is exempt
dbGetQuery(mydb,"INSERT INTO test2 (score, student) VALUES (NULL, 'A');")
dbGetQuery(mydb,"INSERT INTO test2 (score, student) VALUES (NULL, 'A');")

#EXERCISE 5
#Recreate one of your toy data tables with the constraint that for one of the integer variablesthe default value will be zero. Test your table by inserting some empty values. Display your new tables. Then delete your table.  

```
```{r}
#EXERCISE 5
#Recreate one of your toy data tables with the constraint that for one of the integer variablesthe default value will be zero. Test your table by inserting some empty values. Display your new tables. Then delete your table.
#drop the original table
dbGetQuery(mydb, "DROP TABLE IF EXISTS SchoolRankings;" )
#create the table and set default value of integer variables as 0
dbGetQuery(mydb, "CREATE TABLE SchoolRankings (
           SchoolName TEXT,
           SchoolStateRanking INTEGER DEFAULT 0,
           SchoolNationalRanking INTEGER DEFAULT 0
);")
#check if the table successfully create 
dbListTables(mydb)
#insert new empty values 
#dbGetQuery(mydb,"INSERT INTO test2 (score, student) VALUES ('1', 'A');")
dbGetQuery(mydb, "INSERT INTO SchoolRankings (SchoolName) Values ('a');")
dbGetQuery(mydb, "INSERT INTO SchoolRankings (SchoolName,SchoolNationalRanking) Values ('b',3);")

#display new table and check if the new data was successffuly inserted 
dbGetQuery(mydb, "SELECT * FROM SchoolRankings;")
#Drop the table and check 
dbGetQuery(mydb, "DROP TABLE IF EXISTS SchoolRankings;")
dbListTables(mydb)

```




# Adding a column with a default value
```{r}
#Add a column with default value 1
#dbGetQuery(mydb,"SELECT * FROM studentAssessment LIMIT 20;")

dbGetQuery(mydb, "ALTER TABLE studentAssessment ADD email INTEGER DEFAULT 1 ")#The ALTER TABLE statement is used to add, delete, or modify columns in an existing table.

dbGetQuery(mydb, "SELECT * FROM studentAssessment LIMIT 10;")

#Delete a column
dbGetQuery(mydb, "ALTER TABLE studentAssessment DROP COLUMN email;")


#EXERCISE 6
#Add a column to one of your toy data tables with a default value of 3. Display your new table. Delete this column.
```

```{r}
#EXERCISE 6
#Add a column to one of your toy data tables with a default value of 3. Display your new table. Delete this column.
#add new column for number of courses that a teacher teaches in table NewTeachers and set default value 3
dbGetQuery(mydb, "ALTER TABLE NewTeachers ADD NumofCourses INTEGER DEFAULT 3;")
#display the new column and delete the column 
dbGetQuery(mydb,"SELECT * FROM NewTeachers;")
dbGetQuery(mydb,"ALTER TABLE NewTeachers DROP COLUMN NumofCourses;")
```

# ID Columns

```{r}
#Auto-increment allows a unique number to be generated automatically when a new record is inserted into a table.
#Often this is the primary key field that we would like to be created automatically every time a new record is inserted.
dbGetQuery(mydb,"CREATE TABLE test3 (
  id INTEGER AUTO_INCREMENT PRIMARY KEY, #Not standard syntax
  score INTEGER, 
  student TEXT
  );")

dbGetQuery(mydb,"INSERT INTO test3 (score, student) VALUES (1, 'A');")
dbGetQuery(mydb,"INSERT INTO test3 (score, student) VALUES (5, 'B');")

dbGetQuery(mydb, "SELECT * FROM test3;")

dbGetQuery(mydb, "DROP TABLE IF EXISTS test3;")

#EXERCISE 7
#Create a new table with four variables and a primary key that is a sequential id value.

```
```{r}
#EXERCISE 7
#Create a new table with four variables and a primary key that is a sequential id value.
#dbListTables(mydb)
dbGetQuery(mydb, "DROP TABLE IF EXISTS Schools;")
dbGetQuery(mydb,"CREATE TABLE Schools (
           Schoolid INTEGER AUTO_INCREMENT PRIMARY KEY,
           SchoolName TEXT,
           SchoolNationalRanking INTEGER DEFAULT 3,
           SchoolStateRanking INTEGER DEFAULT 1
           
);")
#insert some new values 
dbGetQuery(mydb,"INSERT INTO Schools(SchoolName) VALUES ('a');")
dbGetQuery(mydb,"INSERT INTO Schools(SchoolName) VALUES ('b');")
dbGetQuery(mydb,"INSERT INTO Schools(SchoolName) VALUES ('c');")
dbGetQuery(mydb,"INSERT INTO Schools(SchoolName) VALUES ('d');")
#display the table
dbGetQuery(mydb, "SELECT * FROM Schools;")

```


## Filtering (WHERE)
```{r}
#dbGetQuery(mydb,"SELECT * FROM studentAssessment LIMIT 10;")
dbGetQuery(mydb, "SELECT id_student, date_submitted FROM studentAssessment WHERE date_submitted > 550 ORDER BY date_submitted DESC;")

#OR Statement
dbGetQuery(mydb, "SELECT id_student, date_submitted FROM studentAssessment WHERE date_submitted > 550 OR date_submitted < 2 ORDER BY date_submitted DESC;")

#AND Statement
dbGetQuery(mydb, "SELECT id_student, date_submitted FROM studentAssessment WHERE date_submitted > 550 AND id_student = 325750 ORDER BY date_submitted DESC;")

#LIKE
#The LIKE operator is used in a WHERE clause to search for a specified pattern in a column.

#There are two wildcards often used in conjunction with the LIKE operator:
#% - The percent sign represents zero, one, or multiple characters
#_ - The underscore represents a single character

dbGetQuery(mydb, "SELECT id_student, gender, region FROM studentInfo WHERE region LIKE '%Region%';")

#Begin with 'Region'
dbGetQuery(mydb, "SELECT id_student, gender, region FROM studentInfo WHERE region LIKE 'Region%';")

#End with 'Region'
dbGetQuery(mydb, "SELECT id_student, gender, region FROM studentInfo WHERE region LIKE '%Region';")

#'c' is the second letter
dbGetQuery(mydb, "SELECT id_student, gender, region FROM studentInfo WHERE region LIKE '_c%';")

#IN
#The IN operator allows you to specify multiple values in a WHERE clause.
#The IN operator is a shorthand for multiple OR conditions.
dbGetQuery(mydb, "SELECT id_student, gender, region FROM studentInfo WHERE region IN ('Wales','Ireland');")

#EXERCISE 8
#Query one of your original toy data tables, for two different conditions. 

```
```{r}
#EXERCISE 8
#Query one of your original toy data tables, for two different conditions
#dbListTables(mydb)
#dbGetQuery(mydb, "SELECT * FROM NewTeachers;")
#display the information of teachers named as Bob 
dbGetQuery(mydb, "SELECT * FROM NewTeachers WHERE TeacherFirstName LIKE '%Bob%';")
#display the teachers names where teachers first names have letter "o" in the second place
dbGetQuery(mydb,"SELECT TeacherFirstName, TeacherLastName FROM NewTeachers WHERE TeacherFirstName LIKE '_o%';")
```

## Removing Duplicates
```{r}
#dbGetQuery(mydb, "SELECT * FROM studentInfo;")
dbGetQuery(mydb, "SELECT DISTINCT region FROM studentInfo;")

dbGetQuery(mydb, "SELECT DISTINCT region, gender FROM studentInfo;")

#EXERCISE 9
#Insert a duplicate row into one of your toy data tables. Then query the table without including duplicates.
```
```{r}
#EXERCISE 9
#Insert a duplicate row into one of your toy data tables. Then query the table without including duplicates.
dbGetQuery(mydb, "SELECT * FROM NewTeachers;")
#since the original table already has many duplicated rows so I add two new duplicated rows 
dbGetQuery(mydb, "INSERT INTO NewTeachers VALUES('Emily', 'D',55),('Emily','D',55);")
#if it's set up correctly, there should be 32 rows (no spacd b/t count (*))
dbGetQuery(mydb, "SELECT COUNT(*) FROM NewTeachers;")
#displace unique rows 
dbGetQuery(mydb, "SELECT DISTINCT TeacherFirstName, TeacherLastName, TeacherAge FROM NewTeachers;")
#dbGetQuery(mydb, "SELECT * FROM NewTeachers;")
```



## Conditional Expressions (non-standard)
##The CASE statement goes through conditions and returns a value when the first condition is met (like an IF-THEN-ELSE statement). So, once a condition is true, it will stop reading and return the result. If no conditions are true, it returns the value in the ELSE clause.
##CASE
##    WHEN condition1 THEN result1
##    WHEN condition2 THEN result2
##    WHEN conditionN THEN resultN
##    ELSE result
##END;
If there is no ELSE part and no conditions are true, it returns NULL.
```{r}
dbGetQuery(mydb, "DROP TABLE IF EXISTS booltest")
dbGetQuery(mydb, "CREATE TABLE booltest (a INTEGER, b INTEGER);")
dbGetQuery(mydb, "INSERT INTO booltest VALUES (1, 0);")
dbGetQuery(mydb, "SELECT * FROM booltest;")

dbGetQuery(mydb,"SELECT
  CASE WHEN a THEN 'true' ELSE 'false' END as boolA,
  CASE WHEN b THEN 'true' ELSE 'false' END as boolB
  FROM booltest")

dbGetQuery(mydb,"SELECT
  CASE a WHEN 1 THEN 'true' ELSE 'false' END as boolA,
  CASE b WHEN 1 THEN 'true' ELSE 'false' END as boolB
  FROM booltest")
```

#Relationships (JOIN) - *Slide*
```{r}

#Create two tables with matches and join them

dbGetQuery(mydb, "CREATE TABLE left_table (id INTEGER, description TEXT);")
dbGetQuery(mydb, "CREATE TABLE right_table (id INTEGER, description TEXT);")

dbGetQuery(mydb, "INSERT INTO left_table VALUES ( 1, 'left 01');")
dbGetQuery(mydb, "INSERT INTO left_table VALUES ( 2, 'left 02');")
dbGetQuery(mydb, "INSERT INTO left_table VALUES ( 3, 'left 03');")
dbGetQuery(mydb, "INSERT INTO left_table VALUES ( 4, 'left 04');")
dbGetQuery(mydb, "INSERT INTO left_table VALUES ( 5, 'left 05');")
dbGetQuery(mydb, "INSERT INTO left_table VALUES ( 6, 'left 06');")
dbGetQuery(mydb, "INSERT INTO left_table VALUES ( 7, 'left 07');")
dbGetQuery(mydb, "INSERT INTO left_table VALUES ( 8, 'left 08');")
dbGetQuery(mydb, "INSERT INTO left_table VALUES ( 9, 'left 09');")

dbGetQuery(mydb, "INSERT INTO right_table VALUES ( 6, 'left 06');")
dbGetQuery(mydb, "INSERT INTO right_table VALUES ( 7, 'left 07');")
dbGetQuery(mydb, "INSERT INTO right_table VALUES ( 8, 'left 08');")
dbGetQuery(mydb, "INSERT INTO right_table VALUES ( 9, 'left 09');")
dbGetQuery(mydb, "INSERT INTO right_table VALUES ( 10, 'left 10');")
dbGetQuery(mydb, "INSERT INTO right_table VALUES ( 11, 'left 11');")
dbGetQuery(mydb, "INSERT INTO right_table VALUES ( 12, 'left 12');")
dbGetQuery(mydb, "INSERT INTO right_table VALUES ( 13, 'left 13');")
dbGetQuery(mydb, "INSERT INTO right_table VALUES ( 14, 'left 14');")

dbGetQuery(mydb, "SELECT * FROM left_table;")
dbGetQuery(mydb, "SELECT * FROM right_table;")

#Rename descriptions column as xx_table and rename table as l or r
dbGetQuery(mydb,"SELECT l.description AS left_table, r.description AS right_table 
           FROM left_table AS l 
           JOIN right_table AS r ON l.id = r.id")

dbGetQuery(mydb,"SELECT l.description AS left_table, r.description AS right_table 
           FROM left_table AS l 
           RIGHT JOIN right_table AS r ON l.id = r.id")

dbGetQuery(mydb,"SELECT l.description AS left_table, r.description AS right_table 
           FROM left_table AS l 
           LEFT JOIN right_table AS r ON l.id = r.id")

#Union
dbGetQuery(mydb, "SELECT * FROM left_table
           UNION
           SELECT * FROM right_table;")


#EXERCISE 10
# Create a common id variable in your two toy data tables. Then join those tables so that your query returns all the values from one table and only those that match from the other. 

```
```{r}
#EXERCISE 10
# Create a common id variable in your two toy data tables. Then join those tables so that your query returns all the values from one table and only those that match from the other. 
dbListTables(mydb)
dbGetQuery(mydb, "SELECT * FROM Schools;")
#drop the unique primary key id
dbGetQuery(mydb, "ALTER TABLE Schools DROP Schoolid;")
dbGetQuery(mydb, "SELECT * FROM NewTeachers;")

#insert a new common column
dbGetQuery(mydb, "ALTER TABLE Schools ADD common_ID INTEGER;")
dbGetQuery(mydb, "ALTER TABLE NewTeachers ADD common_ID INTEGER;")
#check num of rows for prep of inserting new values 
dbGetQuery(mydb, "SELECT COUNT(*) FROM Schools;") #4 rows
dbGetQuery(mydb, "SELECT * FROM Schools;")
dbGetQuery(mydb, "SELECT COUNT(*) FROM NewTeachers;") #32 rows
dbGetQuery(mydb, "SELECT * FROM NewTeachers;")
#insert new rows with common values for two tables 
dbGetQuery(mydb, "INSERT INTO Schools (common_ID) Values (1) ;")
dbGetQuery(mydb, "INSERT INTO Schools (common_ID) Values (2) ;")
dbGetQuery(mydb, "INSERT INTO Schools (common_ID) Values (3) ;")
dbGetQuery(mydb, "INSERT INTO Schools (common_ID) Values (4) ;")
dbGetQuery(mydb, "INSERT INTO Schools (common_ID) Values (5) ;")
dbGetQuery(mydb, "INSERT INTO Schools (common_ID) Values (6) ;")
dbGetQuery(mydb, "INSERT INTO Schools (common_ID) Values (7) ;")
dbGetQuery(mydb, "SELECT * FROM Schools;")

dbGetQuery(mydb, "INSERT INTO NewTeachers (common_ID) Values (1);")
dbGetQuery(mydb, "INSERT INTO NewTeachers (common_ID) Values (2);")
dbGetQuery(mydb, "INSERT INTO NewTeachers (common_ID) Values (3);")
dbGetQuery(mydb, "INSERT INTO NewTeachers (common_ID) Values (4);")
dbGetQuery(mydb, "INSERT INTO NewTeachers (common_ID) Values (5);")
dbGetQuery(mydb, "INSERT INTO NewTeachers (common_ID) Values (6);")
dbGetQuery(mydb, "INSERT INTO NewTeachers (common_ID) Values (7);")
dbGetQuery(mydb, "SELECT * FROM NewTeachers;")


#join 
dbGetQuery(mydb,"SELECT *
           FROM Schools AS l 
           JOIN NewTeachers AS r ON l.common_ID = r.common_ID")
# right join 
#it's very messy 
dbGetQuery(mydb,"SELECT *
           FROM Schools AS l 
           RIGHT JOIN NewTeachers AS r ON l.common_ID = r.common_ID")
#union 
dbGetQuery(mydb, "SELECT * FROM Schools
           UNION
           SELECT * FROM NewTeachers;")

```

```{r}
#Now disconnect from your database
dbDisconnect(mydb)

#Then retunr to your AWS console and:

#1. Click on "Actions" and then "Stop"
#2. Do NOT make a snapshot
#3 Click on "Actions" again and click "Delete"
#4. Unclick "Make a final snapshot"
#5. Clicl "I acknowledge that upon instance deletion, automated backups, including system snapshots and point-in-time recovery, will no longer be available."
#6. Type "delete me" into the field

#Failure to follow these steps could result in charges to your credit card.


```

